outline!

### 2016-05-13 paper.tex
paper: working on outline
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

\documentclass[10pt,twocolumn]{article}
\usepackage{times}
\usepackage{multirow}
\usepackage{graphicx,grffile}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}
\usepackage[latin1]{inputenc}

@@

\title{title}

@@

\begin{abstract}
abstract
\end{abstract}

@@

@@

\section{Recurring Problem}

Lets motivate here. Showing the need for programmability and some 
declarative interface.

Describe corfu and object class interfaces.

Describe CORFU briefly, and then show the physical design challenge. The
performance decision changes depending on the mix of hardware and software
versions and configurations.

Use state diagram to show how this works for corfu.

then show physical design crazy ness as we go to different
versions and/or different hardware.

Here we create a programming model for the objects

Show how the corfu interface is defined in this model.

\section{Storage System Programmability}

\begin{figure}[ht]
  \centering
    \includegraphics[width=0.45\textwidth]{experiments/objclass-dev/output.png}
    \caption{
[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/objclass-dev/visualize.ipynb}{source}]
Growth of officially supported, custom object interfaces in RADOS over 6
years. A \emph{method} is a specific object interface and a \emph{class} is a
logical grouping of methods.
}
\label{fig:objclass-dev}
\end{figure}

When application goals are not met by a storage system the most common reaction
is to design a workaround. Workarounds roughly fall into one of two categories:
so called ``bolt-on'' services that introduce a 3rd party system (e.g. a
metadata service), or expanded application responsibility in the form of data
management (e.g. a new data layout).

\section{Consolidated Section}

Up until this point we have introduced Ceph, CORFU, the general problems
associated with programmability, described ZLog as being an instantiation of
the CORFU interface on top of Ceph.

Template for the CORFU object class interface. Maybe a simple state machine.

Describe indexing and protocol enforcement.
Differences between built-in interfaces and custom cls interfaces.

\section{ZLog Physical Design}

In this section we examine the design space for implementing the CORFU
protocol in the RADOS storage system. Table \ref{t:init-ds} shows the entire
design space. A description of each design parameter follows:

{\bf Mapping strategy.} The method by which a log entry---identified by its
logical position---is addressed within RADOS is referred to as the mapping
strategy. A 1:1 strategy stores each log entry in a RADOS object with a
distinct name (e.g. ``mylog.pos443''), and an N:1 strategy stripes the log
positions across a set of objects (e.g.  round-robin). While we do consider
both strategies in this paper, a 1:1 strategy is attractive because it allows
a design in which clients can directly address log positions by constructing
the correct object name.

{\bf Storage interface.} entry data may be small or large. its primary
storage location is important. kv, bs. we further sub-divide bs into
write and append. each strategy is compatible with fixed size entries
except for the write interface.

{\bf Logical addressing.}

\subsection{Storage Interface Selection}

Figure \ref{f:librados-sweep} shows the single-node write-only I/O throughput
for each of the points in the design space defined in Table \ref{t:init-ds}.
The results reveal poor relative performance using both the key-value storage
interface, as well as either of the 1:1 strategies---which incur overhead by
creating a new object for each log entry. Both of the N:1 strategies using the
bytestream I/O interface outperform all of the other approaches by over 2x
throughput, but otherwise have nearly identical performance.

To examine the read performance of the points in the design space we generated
a dataset using each of the write workloads, and then issued random reads
across the entire dataset. Figure \ref{f:librados-rand-read} shows the read
throughput for each of the design strategies. The results show that an N:1
mapping using bytestream interface performs best. Interestingly, the 1:1
mapping strategy using the bytestream interface exhibits good read performance
in comparison to writes using the same strategy.

\begin{table}
\begin{tabular}{|l|l|l|}
\hline
Category & Specialization & Methods \\ \hline
\multirow{2}{*}{Locking} & Shared & \multirow{2}{*}{6} \\
                         & Exclusive & \\ \hline
\multirow{3}{*}{Logging} & Replica & 3 \\
                         & State & 4 \\
                         & Timestamped & 4 \\ \hline
Garbage Collection & Reference Counting & 4 \\ \hline
\multirow{4}{*}{Metadata Management} & RBD & 37 \\
 & RGW & 27 \\
 & User & 5 \\
 & Version & 5 \\ \hline
\end{tabular}
\caption{A variety of RADOS object storage classes exist that expose reusable interfaces to applications.}
\label{tab:objclasses}
\end{table}

@@

  \path (A) edge        node [left] {R,W,F} (B)
            edge        node {T}     (C)
        (B) edge        node {F}     (C)
            edge        node {R,W}   (D)
        (C) edge        node {F,T}   (E)
        (D) edge        node {W}     (E)
            edge        node {R}     (F)
        (E) edge        node {F,T,W} (F);
\end{tikzpicture}
\caption{State transition diagram for read ($R$), write ($W$), fill ($F$), and
trim ($T$) CORFU operaitons. The states epoch guard ($EG$), check position ($CP$),
and update state ($US$) access metadata. The I/O performs a log entry read or
write, and garbage collection ($GC$) marks entries for reclamation.}
\end{figure}

\begin{table}
\begin{tabular}{ | l | l | l | l | l |}
\hline
Map & I/O & Entry Size & Addressing & Metadata \\ \hline
\multirow{3}{*}{1:1} & KV  & Flex     & Ceph      & KV/BS \\ \cline{2-5}
                     & BS  & Flex     & Ceph/VFS  & KV/BS \\ \hline
\multirow{4}{*}{N:1} & KV  & Flex     & \multicolumn{2}{|c|}{KV/BS} \\ \cline{2-5}
                     & WR  & Fixed    & VFS       & KV/BS \\ \cline{2-5}
                     & AP  & Flex     & KV/BS     & KV/BS \\
\hline
\end{tabular}
\caption{The high-level design space of mapping CORFU log entry storage onto
the RADOS object storage system.}
\label{t:init-ds}
\end{table}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{experiments/librados-sweep/output.soft.reset.2hr.png}
  \caption{
[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/librados-sweep/visualize.ipynb}{source}]
Throughput (IOPS) of 1K writes to a single OSD using the standard I/O
interfaces in various configurations. The best performance is achieved using
the byte stream interface and a N:1 mapping strategy.
}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{experiments/basic-cls-rand-read/output.read.60min.png}
  \caption{
[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/basic-cls-rand-read/visualize.ipynb}{source}]
Throughput of 1K random reads to a single OSD using the standard I/O
interfaces in various configurations. The best performance is achieved using
the byte stream interface and a N:1 mapping strategy.
}
\end{figure}

As a first approximation these results show that when optimizing for log
append and random read throughput the design space should be refined by
limiting solutions to architectures based on a N:1 mapping strategy using the
bytestream I/O interface. But what these microbenchmarks explicitly omit are
the overheads associated with metadata management such as enforcing write-once
semantics or using an index to map a log position to a physical offset within
an object. In the next section we

\subsection{Metadata Management}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
  \caption{
[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/basic-cls-overhead/visualize.ipynb}{source}]
}
\end{figure}

Many variations of design here. What is the easiest way to 
winnow down the design space.

NEXT GRAPH A

This is an easy interface to simulate.
- Check Epoch (omap, stream header)
- this graph also has the APPEND raw through for comparison
- adds NEW OSD OP TO AVOID CLS OVERHEAD DISCUSSION FOR PAPER
- 3 runs in this graph, no need to write interface
- short runs to verify and then do all long runs togther with next graph

Both protocol enforcement and logical address translation
represent a larger design space involving index design. However
since in either case all state must be stored in the omap and
in the stream header, testing the overhead of epoch enforcement
provides a strict upper bound on performance.

- Protocol Enforcement
- Logical Translation (only append case)

Memory Cache

NEXT GRAPH B (or combined with A)
Use in-memory cache to implement the epoch check and show that we can match
the raw performance even with this added functionality.
This will show the content from graph A for comparison. Look
we match the raw performance.

NEXT GRAPH B or C (probably just two).. bar chart?
 Final consolidation of average throughputs of all the different
methods
 THIS INCLUDES THE FULL IMPL OF ZLOG APPEND using the in-memory cache

We solve the slow down of the epoch by storing the value in memory. Here is
the new performance.  We are able to with the epoch check match the raw
performance.

We know that we cannot use kv/bs for logicla address translation or protocol
enforcement without a performance hit so there is no point in trying. instead,
we generalize this memory cache hting used for epoch checks as a
new interface and use this interface to implement the full zlog
stack and show the performance.

Overall corfu is relatively simple abtraction, yet requires significant
work in making performant. Further, the decision may change over time.

We will probably have room in the paper to discuss the design of the
index structure, like collapsing a bit map and taking advantage of
the implicit strides in the offsets, etc... but we can probably move
that to a later section to keep this presentation concise. We should
probably mention this aspect here, but overall that becomes a ortholognal
optimization.

%# Abstract Model

%# Sequencer Creation

>>>>

\documentclass[10pt,twocolumn]{article}
\usepackage{times}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{graphicx,grffile}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}
\usepackage[latin1]{inputenc}

@@

\title{outline!}

@@

\begin{abstract}
\end{abstract}

@@

\begin{figure}[ht]
  \centering
    \includegraphics[width=0.48\textwidth]{experiments/objclass-dev/output.png}
    \caption{
[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/objclass-dev/visualize.ipynb}{source}]
Growth of officially supported, custom object interfaces in RADOS over 6
years. A \emph{method} is a specific object interface and a \emph{class} is a
logical grouping of methods.
}
\label{fig:objclass-dev}
\end{figure}

@@

\section{Background}

In this section we briefly describe Ceph and object-based interface
development. We will use CORFU as a motivating example of an interface, and
then show the challenges with the current low-level mechanisms for defining
interfaces.

\subsection{Domain-specific Object Interfaces}

Ceph allows the construction of domain specific interfaces. There are many
different object interfaces in production today. An example of the spectrum of
interface types is shown in Table~\ref{tab:objclass-cats}. Object classes are
developed as a transactional composition of native storage interfaces.

\begin{table}
\begin{tabular}{|l|l|l|}
\hline
Category & Specialization & Methods \\ \hline
\multirow{2}{*}{Locking} & Shared & \multirow{2}{*}{6} \\
                         & Exclusive & \\ \hline
\multirow{3}{*}{Logging} & Replica & 3 \\
                         & State & 4 \\
                         & Timestamped & 4 \\ \hline
GC & Reference Counting & 4 \\ \hline
\multirow{4}{*}{Metadata} & RBD & 37 \\
 & RGW & 27 \\
 & User & 5 \\
 & Version & 5 \\ \hline
\end{tabular}
\caption{A variety of RADOS object storage classes exist that expose reusable
interfaces to applications.}
\label{tab:objclass-cats}
\end{table}

\subsection{Motivating Example: CORFU Log}

Short description of CORFU system~\cite{balakrishnan:nsdi12}. It's a log!
While CORFU relies on custom interfaces built into flash devices, an
implementation in Ceph relies upon custom device interfaces to enforce its
protocol.

The state-machine shown in Figure~\ref{fig:corfu-sm} shows the composition of
actions for each component of the CORFU interface. For example, each
operations begins with an \emph{epoch guard}, and a \emph{write} operation
will consult an index and perform a bulk I/O operation.  The primary concern
when mapping an interface onto Ceph is deciding how each native interface is
used in the interface composition.

@@

  \path (A) edge        node [left] {R,W,F} (B)
            edge        node {T}     (C)
        (B) edge        node {F}     (C)
            edge        node {R,W}   (D)
        (C) edge        node {F,T}   (E)
        (D) edge        node {W}     (E)
            edge        node {R}     (F)
        (E) edge        node {F,T,W} (F);
\end{tikzpicture}
\caption{State transition diagram for read ($R$), write ($W$), fill ($F$), and
trim ($T$) CORFU operaitons. The states epoch guard ($EG$), check position ($CP$),
and update state ($US$) access metadata. The I/O performs a log entry read or
write, and garbage collection ($GC$) marks entries for reclamation.}
\label{fig:corfu-sm}
\end{figure}

\subsection{Physical Design}

There is a finite set of strategies for mapping the CORFU interfaces onto the
current set of native I/O interfaces used to construct object interfaces.
Table~\ref{tab:pd-map} shows the design space for mapping the CORFU interfaces
onto Ceph. In order to select the best mapping we performed a parameter sweet
over the design space.

\begin{figure*}[!ht]
    \centering
    \begin{subfigure}{.6\columnwidth}
        \includegraphics[width=\columnwidth]{experiments/librados-sweep/output.soft.reset.2hr.png}
        \caption{a}
        \label{fig:vanilla-io-diff}
    \end{subfigure}\hfill
    \begin{subfigure}{.6\columnwidth}
        \includegraphics[width=\columnwidth]{experiments/basic-cls-rand-read/output.read.60min.png}
        \caption{b}
    \end{subfigure}\hfill
    \begin{subfigure}{.6\columnwidth}
        \includegraphics[width=\columnwidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
        \caption{c}
    \end{subfigure}\hfill
    \caption{Shown here are the graphs and such that demonstrate that the same
        physical design choices are not the same between differing version of
    Ceph even on the same hardware.}
\end{figure*}

Figure~\ref{fig:vanilla-io-diff} shows the result of a parameter sweep which
clearly shows that an N-1 mapping based on the bytestream interface provides
superiour performance. However, the other two graphs show the same interfaces
running on an old version of Ceph that show the same decision would not have
been the optimal choice {\bf \emph{note: for the outline these are not the
real graphs we'll be using}}.

\begin{table}
\begin{tabular}{ | l | l | l | l | l |}
\hline
Map & I/O & Entry Size & Addressing & Metadata \\ \hline
\multirow{3}{*}{1:1} & KV  & Flex     & Ceph      & KV/BS \\ \cline{2-5}
                     & BS  & Flex     & Ceph/VFS  & KV/BS \\ \hline
\multirow{4}{*}{N:1} & KV  & Flex     & \multicolumn{2}{|c|}{KV/BS} \\ \cline{2-5}
                     & WR  & Fixed    & VFS       & KV/BS \\ \cline{2-5}
                     & AP  & Flex     & KV/BS     & KV/BS \\
\hline
\end{tabular}
\caption{The high-level design space of mapping CORFU log entry storage onto
the RADOS object storage system.}
\label{tab:pd-map}
\end{table}

In this section we are going to be describing our way of creating interfaces.

Here we show the derivation of two other interfaces that are in production in
Ceph today to demonstrate the generality of our interface.

\section{Related Work}

\section{Conclusion}

%In this section we examine the design space for implementing the CORFU
%protocol in the RADOS storage system. Table \ref{t:init-ds} shows the entire
%design space. A description of each design parameter follows:
%
%{\bf Mapping strategy.} The method by which a log entry---identified by its
%logical position---is addressed within RADOS is referred to as the mapping
%strategy. A 1:1 strategy stores each log entry in a RADOS object with a
%distinct name (e.g. ``mylog.pos443''), and an N:1 strategy stripes the log
%positions across a set of objects (e.g.  round-robin). While we do consider
%both strategies in this paper, a 1:1 strategy is attractive because it allows
%a design in which clients can directly address log positions by constructing
%the correct object name.
%
%{\bf Storage interface.} entry data may be small or large. its primary
%storage location is important. kv, bs. we further sub-divide bs into
%write and append. each strategy is compatible with fixed size entries
%except for the write interface.
%
%{\bf Logical addressing.}

%\subsection{Storage Interface Selection}
%
%Figure \ref{f:librados-sweep} shows the single-node write-only I/O throughput
%for each of the points in the design space defined in Table \ref{t:init-ds}.
%The results reveal poor relative performance using both the key-value storage
%interface, as well as either of the 1:1 strategies---which incur overhead by
%creating a new object for each log entry. Both of the N:1 strategies using the
%bytestream I/O interface outperform all of the other approaches by over 2x
%throughput, but otherwise have nearly identical performance.
%
%To examine the read performance of the points in the design space we generated
%a dataset using each of the write workloads, and then issued random reads
%across the entire dataset. Figure \ref{f:librados-rand-read} shows the read
%throughput for each of the design strategies. The results show that an N:1
%mapping using bytestream interface performs best. Interestingly, the 1:1
%mapping strategy using the bytestream interface exhibits good read performance
%in comparison to writes using the same strategy.

%\begin{figure}[h]
%  \centering
%  \includegraphics[width=0.45\textwidth]{experiments/librados-sweep/output.soft.reset.2hr.png}
%  \caption{
%[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/librados-sweep/visualize.ipynb}{source}]
%Throughput (IOPS) of 1K writes to a single OSD using the standard I/O
%interfaces in various configurations. The best performance is achieved using
%the byte stream interface and a N:1 mapping strategy.
%}
%\end{figure}
%
%\begin{figure}[h]
%  \centering
%  \includegraphics[width=0.45\textwidth]{experiments/basic-cls-rand-read/output.read.60min.png}
%  \caption{
%[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/basic-cls-rand-read/visualize.ipynb}{source}]
%Throughput of 1K random reads to a single OSD using the standard I/O
%interfaces in various configurations. The best performance is achieved using
%the byte stream interface and a N:1 mapping strategy.
%}
%\end{figure}
%
%\begin{figure}[h]
%  \centering
%  \includegraphics[width=0.45\textwidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
%  \caption{
%[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/basic-cls-overhead/visualize.ipynb}{source}]
%}
%\end{figure}

%As a first approximation these results show that when optimizing for log
%append and random read throughput the design space should be refined by
%limiting solutions to architectures based on a N:1 mapping strategy using the
%bytestream I/O interface. But what these microbenchmarks explicitly omit are
%the overheads associated with metadata management such as enforcing write-once
%semantics or using an index to map a log position to a physical offset within
%an object. In the next section we
